Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-11-02 ms_pre_post.Rmd
A couple more sentences in the last paragraph; fix some spelling

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and monandry is common, evolution of post-copulatory inbreeding avoidance is especially unlikely. For example, females of the gregarious cockroach *Blatella germanica* (L.) avoid inbreeding and are typically monandrous [@Lihoreau2007]; while occurrence of post-copulatory inbreeding avoidance in *B. germanica* has not, to our knowledge, been tested, we predict that such inbreeding avoidance is highly unlikely. Similarly, females of the solitary parasitoid wasp *Venturia canescens* are monandrous, and appear to avoid mating with siblings [@Metzger2010; @Metzger2010a]. We suggest that post-copulatory inbreeding avoidance in *V. canescens* would be highly unlikely to evolve.

@@

Some broad predictions are possible given the verbal theory that we have developed and the populations that we have simulated. For example, given that post-copulatory inbreeding avoidance is useless for monandrous females and most effective for polyandrous females, while pre-copulatory inbreeding avoidance is perhaps most effective for monandrous females and less effective for polyandrous females, it would be interesting to test whether or not occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandory. To address such a question, more work is needed to identify the degree to which females of different species engage in polyandrous and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

>>>>

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and monandry is common, evolution of post-copulatory inbreeding avoidance is especially unlikely. For example, females of the gregarious cockroach *Blatella germanica* (L.) avoid inbreeding and are typically monandrous [@Lihoreau2007]; while occurrence of post-copulatory inbreeding avoidance in *B. germanica* has not, to our knowledge, been tested, we predict that such inbreeding avoidance is highly unlikely. Similarly, females of the solitary parasitoid wasp *Venturia canescens* are monandrous, and appear to avoid mating with siblings [@Metzger2010; @Metzger2010a]. We hypothesise that post-copulatory inbreeding avoidance in *V. canescens* would be highly unlikely to evolve.

@@

Some broad predictions are possible given the verbal theory that we have developed and the populations that we have simulated. For example, given that post-copulatory inbreeding avoidance is useless for monandrous females and most effective for polyandrous females, while pre-copulatory inbreeding avoidance is perhaps most effective for monandrous females and less effective for polyandrous females, it would be interesting to test whether or not occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. In addition to polyandry making post-copulatory inbreeding avoidance relevant, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself. Hence, well-grounded biological hypotheses might be derived for the relationship between female multiple mating and mode of inbreeding avoidance. To test such hypotheses, more work is needed to identify the degree to which females of different species engage in polyandrous and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.
